---
title: "Prediction Assignment - Practical Machine Learning Coursera"
author: "Mehdi"
date: "9 January 2018"
output: html_document
---

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here (see the section on the Weight Lifting Exercise Dataset).

# Loading and Processing the Raw Data
The data for this project come from this source

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r , warning=FALSE , message=FALSE }
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(gbm)
library(plyr)

# set the URL for the download
#UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
#UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
dt_training <- read.csv("pml-training.csv")
dt_testing  <- read.csv("pml-testing.csv")

features <- names(dt_testing[,colSums(is.na(dt_testing)) == 0])[8:59]

# Only use features used in testing cases.
dt_training <- dt_training[,c(features,"classe")]
dt_testing <- dt_testing[,c(features,"problem_id")]
dim(dt_training); 
dim(dt_testing);

set.seed(12345)
inTrain <- createDataPartition(dt_training$classe, p=0.6, list=FALSE)
training <- dt_training[inTrain,]
testing <- dt_training[-inTrain,]
dim(training); 
dim(testing);
```

# Building the Decision Tree Model
Using Decision Tree, we shouldn't expect the accuracy to be high. In fact, anything around 80% would be acceptable.

```{r}
modFitDT <- rpart(classe ~ ., data = training, method="class")
fancyRpartPlot(modFitDT)
```


# Predicting with the Decision Tree Model
```{r}
set.seed(12345)

prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$class)
```

#Building the Random Forest Model
Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.
```{r}
set.seed(12345)
modFitRF <- randomForest(classe ~ ., data = training, ntree = 1000)
```

# Predicting on the Testing Data (pml-testing.csv)
```{r}
predictionDT <- predict(modFitDT, dt_testing, type = "class")
predictionDT
```

#Random Forest Prediction
```{r}
predictionRF <- predict(modFitRF, dt_testing, type = "class")
predictionRF
```